---
title: "Replication of Alt et al. (2014) with Corrected Election Timing Data"
author: "Christopher Gandrud"
date: "25 September 2014"
output: pdf_document
bibliography: rep.bib
---

```{r, include=FALSE}
# Uses output files created in Stata 12.1 in data_source/sup003_updated.do

# Set working directory, change as needed
setwd('/git_repositories/Alt_et_al_2014_Replication/')

# Load packages
library(magrittr)
library(DataCombine)

# Load individual tables
files <- list.files('Stata_temp/')

for (i in files){
    if (i == files[[1]]) {
        comb <- foreign::read.dta(paste0('Stata_temp/', i))
    }
    else if (i != files[[1]]) {
        temp <- foreign::read.dta(paste0('Stata_temp/', i))
        temp <- temp[1:94, ] # drop ancillary rows
        comb <- data.frame(comb, temp)  
    }
}

#### Create individual comparative tables ####

# Clean up variable names
from <- c(comb[grep('*_coef', comb), 1][1:9], '^.*?_stderr')
to <- c('Electoral Term', 'Electoral term x Budget transparency',
              'SGP', 'SGP x Budget transparency', 'Slump',
              'Slump x Budget transparency', 'Boom',
              'Boom x Budget transparency','Banking crisis', '')
replacements <- data.frame(from, to)

table_clean <- function(x){
    names(x) <- c('Dependent variable', 'Original', 'Corrected')
    x <- FindReplace(x, Var = 'Dependent variable', replaceData = replacements,
                     exact = F)
    return(x)
}

balance <- comb[1:18, c(1, 2, 4)] %>% table_clean()
debt_change <- comb[1:18, c(1, 6, 8)] %>% table_clean()
sfa <- comb[1:18, c(1, 10, 12)] %>% table_clean()
sfa_noGreece <- comb[1:18, c(1, 14, 16)] %>% table_clean()
net_acq <- comb[1:18, c(1, 18, 20)] %>% table_clean()
net_liab <- comb[1:18, c(1, 22, 24)] %>% table_clean()
```

> This note replicates [@Alt2014] using using corrected election timing data.
A key proposition in this paper--that politicians used 'fiscal gimmickry conditional
on the level of transparency--was tested with election timing data from
[@DPI2012, updated 2012]. This data, however, has serious reliability and validity
issues. The purpose of this note is to discover if the main findings in [@Alt2014]
are robust to a corrected election timing variable. Overall the results are robust,
though the estimated effects' magnitude and statistical significance are lower than
originally reported.

[@Alt2014] recently made an important contribution to the study of fiscal gimmickry

## Replication results

The following tables replicate the main result from [@Alt2014].

### Balance

```{r, echo=FALSE, results='asis'}
knitr::kable(balance)
```

### Debt change

```{r, echo=FALSE, results='asis'}
knitr::kable(debt_change)
```

### SFA

```{r, echo=FALSE, results='asis'}
knitr::kable(sfa)
```

### SFA (excluding Greece)

```{r, echo=FALSE, results='asis'}
knitr::kable(sfa_noGreece)
```

### Net acquisition (+) of shares and other equity

```{r, echo=FALSE, results='asis'}
knitr::kable(net_acq)
```

### Net incurrence (-) of other liabilities

```{r, echo=FALSE, results='asis'}
knitr::kable(net_liab)
```


## References
